Supplementary Information 1

### 2025-04-11 index.Rmd
friday!

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
library(openxlsx)
library(ggthemes)
library(forecast)
library(tidyr)
library(mixR)
library(readxl)
library(tidyverse)
library(ggridges)
```

@@

@@

@@

```
Source of data inpit

@@

p1 <- ggplot(dat1_std, aes(x = year, y = index_std, group = index, color = category)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~index, scales = "free_y", ncol = 5) +  # Ajusta ncol para obtener 3 filas
  scale_color_manual(values = c("Fishery" = '#1b9e77', 
                                "Survey" = '#d95f02',
                                "Predator" = '#7570b3')) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "", x = "", y = "Standard index", color = "Category")
p1
```

@@

```{r }
par(mfrow=c(1,2), mar=c(5,4,2,1))
SSplotData(base.model1.1, 
           subplots = 1,
           pheight = 15)

@@

```{r fig.height=3, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1)) 
SSplotTimeseries(base.model1.1,
                 subplot = 1)
SSplotTimeseries(base.model1.1,
                 subplot = 9)
SSplotTimeseries(base.model1.1,
                 subplot = 7)
```

@@

```{r fig.height=3, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.2,
                 subplot = 1)
SSplotTimeseries(base.model1.2,
                 subplot = 9)
SSplotTimeseries(base.model1.2,
                 subplot = 7)
```

@@

```{r fig.height=3, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.3,
                 subplot = 1)
SSplotTimeseries(base.model1.3,
                 subplot = 9)
SSplotTimeseries(base.model1.3,
                 subplot = 7)
```

@@

```{r fig.height=3, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.4,
                 subplot = 1)
SSplotTimeseries(base.model1.4,
                 subplot = 9)
SSplotTimeseries(base.model1.4,
                 subplot = 7)
```
Selectivity

```{r}
par(mfrow=c(2,2), mar=c(5,4,1,1))
SSplotSelex(base.model1.1,
            subplots = 1)
mtext("s1.1", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.2,
            subplots = 1)
mtext("s1.2", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.3,
            subplots = 1)
mtext("s1.3", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.4,
            subplots = 1)
mtext("s1.4", side=3, line=0.5, cex=1.2)
```
```{r}
SSplotBiology(base.model1.1,
            subplots = 24)
```

@@

```{r}
par(mfrow=c(2,2), mar=c(5,4,1,1))
SSplotBiology(base.model1.1,
            subplots = 21)
mtext("s1.1", side=3, line=0.5, cex=1.2)
SSplotBiology(base.model1.2,
            subplots = 21)
mtext("s1.2", side=3, line=0.5, cex=1.2)
SSplotBiology(base.model1.3,
            subplots = 21)
mtext("s1.3", side=3, line=0.5, cex=1.2)
SSplotBiology(base.model1.4,
            subplots = 21)
mtext("s1.4", side=3, line=0.5, cex=1.2)
```

@@

```{r}
SSplotComps(base.model1.1,
            subplots = 21)
SSplotComps(base.model1.2,
            subplots = 21)
SSplotComps(base.model1.3,
            subplots = 21)
SSplotComps(base.model1.4,
            subplots = 21)
```

Total biomass

@@

ggplot(totalbio, aes(x = Yr, y = Bio_all, color=Serie)) +
  geom_point() +
  geom_smooth(method = "loess",
              span=0.5, 
              se=F)+
  labs(title = "",
       x = "",
       y = "Total biomass (mt)") +
  theme_few() +
  scale_color_viridis_d(option="H",
                        name="Scenarios")+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1))
```

@@

```{r}
df <- base.model1.4$lendbase

df1 <- df %>% 
  dplyr::filter(Pearson<5) %>% 
  dplyr::select(c(1, 6, 19, 16)) %>% 
  dplyr::mutate(Fleet = dplyr::case_when(
    Fleet == 1 ~ "FISHERYBS",
    Fleet == 2 ~ "FISHERYEI",
    Fleet == 3 ~ "FISHERYGS",
    Fleet == 4 ~ "FISHERYJOIN",
    Fleet == 5 ~ "FISHERYSSIW",
    Fleet == 6 ~ "SURVEYBS",
    Fleet == 7 ~ "SURVEYEI",
    Fleet == 8 ~ "SURVEYGS",
    Fleet == 9 ~ "SURVEYJOIN",
    Fleet == 10 ~ "SURVEYSSIW",
    Fleet == 11 ~ "PREDATOR",
    TRUE ~ NA_character_  # Para manejar cualquier valor que no coincida
  ))

# Crea el heatmap
ggplot(df1, aes(x = Yr, y = Bin, fill = Pearson)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", 
                       mid = "white", 
                       high = "red",
                       midpoint = 0) +
  facet_wrap(Fleet~., ncol=5) + 
  labs(title = "",
       x = "",
       y = "Length (cm)",
       fill = "Pearson") +
  theme_few()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1, 
                                   size = 7))

```
hexagon

@@

```{r}
par(mfrow=c(2,4), mar=c(5,4,2,1)) 
SSplotRunstest(base.model1.1,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
```

```{r}
par(mfrow=c(3,3), mar=c(5,4,2,1))
SSplotRunstest(base.model1.2,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
```

```{r}
par(mfrow=c(2,4), mar=c(5,4,2,1))
SSplotRunstest(base.model1.3,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
```

```{r}
par(mfrow=c(3,3), mar=c(5,4,2,1))
SSplotRunstest(base.model1.4,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
```

```{r eval=FALSE}
SSplotRunstest(base.model1.1,
               subplots = "cpue",
               add=T,
               print = TRUE)
SSplotRunstest(base.model1.2,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model1.3,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model1.4,
               subplots = "cpue",
               add=T)
```

@@

```{r}
# Crear un data frame con los valores de RMSE
dfrmse <- data.frame(
  Model1.1 = as.numeric(base.model1.1$index_variance_tuning_check$RMSE),
  Model1.2 = as.numeric(base.model1.2$index_variance_tuning_check$RMSE[1:10]),
  Model1.3 = as.numeric(base.model1.3$index_variance_tuning_check$RMSE[1:10]),
  Model1.4 = as.numeric(base.model1.4$index_variance_tuning_check$RMSE[1:10])

@@

t.test(dfrmse$Model1.1, dfrmse$Model1.2)  # Comparar Model1.1 vs Model1.2
t.test(dfrmse$Model1.1, dfrmse$Model1.3)  # Comparar Model1.1 vs Model1.3
t.test(dfrmse$Model1.2, dfrmse$Model1.3)
t.test(dfrmse$Model1.3, dfrmse$Model1.4)
# ANOVA para ver si hay diferencias globales
anova_rmse <- aov(as.numeric(unlist(dfrmse)) ~ rep(1:4, each=nrow(dfrmse)))
summary(anova_rmse)

# Prueba post-hoc si ANOVA es significativa
#TukeyHSD(anova_rmse)

dfrmse_long <- reshape2::melt(dfrmse)
ggplot(dfrmse_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(width=0.3) +
  geom_jitter(width = 0.01, alpha = 0.3) +
  theme_minimal() +
  scale_fill_manual(name= "Scenario",
                    values = c("Model1.1" = "#ca0020", 
                               "Model1.2" = "#f4a582", 
                               "Model1.3" = "#bababa", 
                               "Model1.4" = "#404040")) +
  labs(title = "", x = "", y = "RMSE")
```
This boxplot compares a summary of the Root Mean Square Error (RMSE) across four different models (s1.1, s1.2, s1.3, and s1.4) used to evaluate recruitment estimates of Antarctic krill. RMSE serves as an indicator of model accuracy, with lower values representing better predictive performance.

@@

```{r}
# First dataset
l <- base.model1.1$recruit$SpawnBio  
m <- base.model1.1$recruit$pred_recr 
yr <- base.model1.1$timeseries$Yr[1:35]

# Second dataset
leco <- base.model1.4$recruit$SpawnBio  
meco <- base.model1.4$recruit$pred_recr 
yreco <- base.model1.4$timeseries$Yr[1:35]

# Beverton-Holt models
bev_holt_model1 <- nls(m ~ (a * l) / (1 + b * l), 
                        start = list(a = max(m), b = 1 / max(l)), 
                        data = data.frame(l, m))

bev_holt_model2 <- nls(meco ~ (a * leco) / (1 + b * leco), 
                        start = list(a = max(meco), b = 1 / max(leco)), 
                        data = data.frame(leco, meco))

# Predictions
l_pred1 <- seq(min(l), max(l), length.out = 100)
m_pred1 <- predict(bev_holt_model1, newdata = data.frame(l = l_pred1))

l_pred2 <- seq(min(leco), max(leco), length.out = 100)
m_pred2 <- predict(bev_holt_model2, newdata = data.frame(leco = l_pred2))

# Create data frames for plotting
data1 <- data.frame(l = l, m = m, yr = yr, Group = "s1.1")
data2 <- data.frame(l = leco, m = meco, yr = yreco, Group = "s1.4")

# Plot
ggplot() +
  geom_point(data = data1, aes(x = l, y = m, shape = Group),
             size = 1.5) +
  geom_text(data = data1, aes(x = l, y = m, label = yr), 
            hjust = 0.5, vjust = -1.5, size = 3) +  
  geom_line(data = data.frame(l = l_pred1, m = m_pred1), 
            aes(x = l, y = m), linewidth = 1,
            col="blue") +
  geom_point(data = data2, aes(x = l, y = m, shape = Group), 
             size = 1.5) +
  geom_text(data = data2, aes(x = l, y = m, label = yr), 
            hjust = 0.5, vjust = 1.5, size = 3) +
  geom_line(data = data.frame(l = l_pred2, m = m_pred2), 
            aes(x = l, y = m), linewidth = 1,
            col="blue") +
  scale_shape_manual(values = c("s1.1" = 5, 
                                "s1.4" = 4)) +
  theme_few() +
  ylim(0, max(m, meco)) +
  labs(x = "SSB", y = "Recruit", 
       title = "", shape = "Model") +
  theme(legend.position = "bottom")

@@

```
### AKL

>>>>

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
library(openxlsx)
library(ggthemes)
library(forecast)
library(tidyr)
library(mixR)
library(readxl)
library(tidyverse)
library(ggridges)
library(ggrepel)
```

@@

The study area corresponds to Subarea 48.1 in the WAP, where most krill fishing activity occurs and where CCAMLR has been working to implement a new management strategy at a reduced spatial scale (Figure \@ref(fig:mapa)). To refine the spatial resolution of krill population dynamics and analysis, we adopted the five strata proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest [@Dornam2021; REF] (Fig 1 S1). While these strata were originally designed for management purposes without considering specific biological or population characteristics, their subdivision enables us to analyze the data at a finer scale. This level of detail facilitates the identification of heterogeneity across various data sources, whether these originate from the fishery or from environmental survey data, providing a more nuanced understanding of krill population dynamics. This spatial segregation of information is incorporated into the stock assessment model, making it a spatially explicit model that accounts for regional variability in population dynamics and external factors.

```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
knitr::include_graphics('Figs/481.png')
```

In this approach we include spatial structure how an implicit approach where different areas are treated as separate fleets. Differences among strata have been identified from various perspectives of krill population dynamics, as well as from the influence of environmental variability within Subarea 48.1 (Figure \@ref(fig:concem)).

```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
knitr::include_graphics('Figs/conceptual.jpeg')
```

@@

```{r path, fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
knitr::include_graphics('Figs/pathmod.png')
```

@@

```
Source of data input

@@

p1 <- ggplot(dat1_std, aes(x = year, y = index_std, group = index, color = category)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~index, scales = "free_y", ncol = 5) +  # Ajusta ncol para obtener 3 filas
  scale_color_manual(values = c("Fishery" = '#1b9e77', 
                                "Survey" = '#d95f02',
                                "Predator" = '#7570b3')) + 
  theme_bw() +
  theme(legend.position = "top",
    axis.text.x = element_text(angle = 90, hjust = 1),
         axis.text.y = element_blank()) +
  labs(title = "", x = "", y = "Standard index", color = "Category")
p1
```

@@

```{r fig.height=8, fig.width=6}
par(mfrow=c(2,1), mar=c(5,4,2,1))
SSplotData(base.model1.1, 
           subplots = 1,
           pheight = 15)

@@

```{r fig.height=2, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1)) 
SSplotTimeseries(base.model1.1,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.1,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.1,
                 subplot = 7,
                 xlab = "")
```

@@

```{r fig.height=2, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.2,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.2,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.2,
                 subplot = 7,
                 xlab = "")
```

@@

```{r fig.height=2, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.3,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.3,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.3,
                 subplot = 7,
                 xlab = "")
```

@@

```{r fig.height=2, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.4,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.4,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.4,
                 subplot = 7,
                 xlab = "")
```
Selectivity

```{r}
par(mfrow=c(2,2), mar=c(5,4,1,1))
SSplotSelex(base.model1.1,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.1", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.2,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.2", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.3,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.3", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.4,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.4", side=3, line=0.5, cex=1.2)
```
```{r}
par(mfrow=c(2,2), mar=c(5,4,1,1))
SSplotBiology(base.model1.1,
            subplots = 1)
SSplotBiology(base.model1.2,
            subplots = 1)
SSplotBiology(base.model1.3,
            subplots = 1)
SSplotBiology(base.model1.4,
            subplots = 1)
```

@@

```{r eval=FALSE}
par(mfrow=c(4,2), mar=c(5,4,2,1))
SSplotBiology(base.model1.1,
              subplots = 21)
SSplotBiology(base.model1.2,
            subplots = 21)
mtext("s1.2", side=3, line=0.5, cex=1.2)
SSplotBiology(base.model1.3,
            subplots = 21)
mtext("s1.3", side=3, line=0.5, cex=1.2)
SSplotBiology(base.model1.4,
            subplots = 21)
mtext("s1.4", side=3, line=0.5, cex=1.2)
```

@@

```{r}
SSplotComps(base.model1.1,
            subplots = 21)
SSplotComps(base.model1.2,
            subplots = 21)
SSplotComps(base.model1.3,
            subplots = 21)
SSplotComps(base.model1.4,
            subplots = 21)
```
Total biomass

@@

ggplot(totalbio, aes(x = Yr, y = Bio_all, color=Serie)) +
  geom_point() +
  geom_smooth(method = "loess",
              span = 0.5,
              se = FALSE) +
  labs(title = "",
       x = "",
       y = "Total biomass (mt)") +
  theme_few() +
  scale_color_manual(name = "",
                     values = c("Without Ecosystem Variables" = "#ca0020",  
                                "W/ Predator" = "#f4a582",  
                                "W/ Environment" = "#bababa",  
                                "With Predator and Ecosystem" = "#404040")) +
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1))

@@

create_heatmap <- function(df) {
  df1 <- df %>%
    dplyr::filter(Pearson < 5) %>%
    dplyr::select(c(1, 6, 19, 16)) %>%
    dplyr::mutate(Fleet = dplyr::case_when(
      Fleet == 1 ~ "FISHERYBS",
      Fleet == 2 ~ "FISHERYEI",
      Fleet == 3 ~ "FISHERYGS",
      Fleet == 4 ~ "FISHERYJOIN",
      Fleet == 5 ~ "FISHERYSSIW",
      Fleet == 6 ~ "SURVEYBS",
      Fleet == 7 ~ "SURVEYEI",
      Fleet == 8 ~ "SURVEYGS",
      Fleet == 9 ~ "SURVEYJOIN",
      Fleet == 10 ~ "SURVEYSSIW",
      Fleet == 11 ~ "PREDATOR",
      TRUE ~ NA_character_
    ))

  ggplot(df1, aes(x = Yr, y = Bin, fill = Pearson)) +
    geom_tile() +
    scale_fill_gradient2(low = "black", 
                         mid = "white", 
                         high = "red", 
                         midpoint = 0) +
    facet_wrap(Fleet~., ncol=6) + 
    labs(title = "", 
         x = "", 
         y = "Length (cm)", 
         fill = "Pearson") + 
    theme_few() + 
    theme(axis.text.x = element_text(angle = 90, 
                                     hjust = 1, 
                                     size = 7),
          legend.position = "bottom")
}

# Crear heatmaps para los diferentes dataframes
heatmap1 <- create_heatmap(base.model1.1$lendbase)
heatmap2 <- create_heatmap(base.model1.2$lendbase)
heatmap3 <- create_heatmap(base.model1.3$lendbase)
heatmap4 <- create_heatmap(base.model1.4$lendbase)

# Combina los plots en un solo gráfico con 8 columnas y 1 fila, compartiendo la leyenda
ggarrange(heatmap1, heatmap2, heatmap3, heatmap4, 
          ncol = 2, 
          nrow = 2, 
          common.legend = TRUE, 
          legend = "bottom")
```

This figure shows the Pearson residuals of predicted length distributions for krill across four modeling scenarios, each incorporating different levels of ecosystem complexity. The residuals are visualized by year and length bin, allowing an assessment of model fit over time and across the size structure of the population. In the "Without Ecosystem Variables" scenario (top left), substantial lack of fit is evident across multiple years and length classes, particularly in the late 1990s and early 2000s. Positive Pearson residuals indicate that the model tends to underestimate frequencies of certain length bins, especially among smaller and mid-sized individuals, suggesting poor representation of recruitment and growth dynamics in the absence of environmental and predator effects. The "With Predator" scenario (top right) shows some improvement in model fit, with slightly more balanced residuals. However, there are still consistent deviations in the mid-size ranges across multiple years, indicating that predation alone does not fully explain observed variability in length composition. In the "With Environment" scenario (bottom left), there is a noticeable reduction in residual magnitude across several years and size classes, implying that incorporating environmental covariates helps align the model more closely with observed size distributions. This is particularly apparent in the early 2000s, where the residuals are closer to zero across most bins. Finally, the "With Predator and Ecosystem" scenario (bottom right) shows the most consistent reduction in Pearson residuals across time and size structure. The distribution of residuals is more homogeneous and centered around zero, indicating that the combined inclusion of predator and environmental drivers leads to the best model fit among the four scenarios.

Overall, the results support the conclusion that ecosystem-informed models better capture the temporal and size-structured dynamics of krill, with the **joint inclusion of predator and environmental variables** yielding the most robust representation of observed length compositions.

@@

```{r}
par(mfrow=c(2,4), mar=c(5,4,2,1)) 
SSplotRunstest(base.model1.1,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
par(mfrow=c(2,5), mar=c(5,4,2,1)) 
SSplotRunstest(base.model1.2,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
par(mfrow=c(2,4), mar=c(5,4,2,1)) 
SSplotRunstest(base.model1.3,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
par(mfrow=c(2,5), mar=c(5,4,2,1)) 
SSplotRunstest(base.model1.4,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
```

```{r}
par(mfrow=c(2,5), mar=c(5,4,2,1))
SSplotRunstest(base.model1.1,
               subplots = "cpue",
               add=T,
               plot = TRUE,
               plotdir = Figs)
par(mfrow=c(2,5), mar=c(5,4,2,1))
SSplotRunstest(base.model1.2,
               subplots = "cpue",
               add=T,
               plot = TRUE,
               plotdir = Figs)
par(mfrow=c(2,5), mar=c(5,4,2,1))
SSplotRunstest(base.model1.3,
               subplots = "cpue",
               add=T,
               plot = TRUE,
               plotdir = Figs)
par(mfrow=c(2,5), mar=c(5,4,2,1))
SSplotRunstest(base.model1.4,
               subplots = "cpue",
               add=T,
               plot = TRUE,
               plotdir = Figs)
```

@@

```{r fig.height=3, fig.width=3}
dfrmse <- data.frame(
  s1.1 = as.numeric(base.model1.1$index_variance_tuning_check$RMSE),
  s1.2 = as.numeric(base.model1.2$index_variance_tuning_check$RMSE[1:10]),
  s1.3 = as.numeric(base.model1.3$index_variance_tuning_check$RMSE[1:10]),
  s1.4 = as.numeric(base.model1.4$index_variance_tuning_check$RMSE[1:10])

@@

t.test(dfrmse$s1.1, dfrmse$s1.2)  # Comparar s1.1 vs s1.2
t.test(dfrmse$s1.1, dfrmse$s1.3)  # Comparar s1.1 vs s1.3
t.test(dfrmse$s1.2, dfrmse$s1.3)
t.test(dfrmse$s1.3, dfrmse$s1.4)
anova_rmse <- aov(as.numeric(unlist(dfrmse)) ~ rep(1:4, each=nrow(dfrmse)))

dfrmse_long <- reshape2::melt(dfrmse)
rmse <-ggplot(dfrmse_long, aes(x = variable, 
                        y = value, 
                        fill= variable)) +
  geom_boxplot(width=0.2,
               alpha=0.8) +
  geom_jitter(width = 0.01, alpha = 0.3) +
  theme_minimal() +
  scale_fill_manual(name= "Scenario",
                    values = c("s1.1" = "#ca0020", 
                               "s1.2" = "#f4a582", 
                               "s1.3" = "#bababa", 
                               "s1.4" = "#404040")) +
  labs(title = "", x = "", y = "RMSE")+
  theme(legend.position = "none")
rmse
```

This boxplot compares a summary of the Root Mean Square Error (RMSE) across four different models (s1.1, s1.2, s1.3, and s1.4) used to evaluate recruitment estimates of Antarctic krill. RMSE serves as an indicator of model accuracy, with lower values representing better predictive performance.

@@

```{r}
# Lista de modelos base
models <- list(
  s1.1 = base.model1.1,
  s1.2 = base.model1.2,
  s1.3 = base.model1.3,
  s1.4 = base.model1.4
)

colors <- c("s1.1" = "#ca0020", "s1.2" = "#f4a582", "s1.3" = "#bababa", "s1.4" = "#404040")
shapes <- c("s1.1" = 5, "s1.2" = 6, "s1.3" = 15, "s1.4" = 4)

plot_data <- list()
curve_data <- list()
label_data <- list()
params <- data.frame(Model = character(), a = numeric(), b = numeric(), stringsAsFactors = FALSE)
for (group in names(models)) {
  model <- models[[group]]
  l <- model$recruit$SpawnBio
  m <- model$recruit$pred_recr
  yr <- model$timeseries$Yr[1:35]

  # Filtrar valores válidos
  valid <- which(is.finite(l) & is.finite(m) & l > 0 & m > 0)
  l <- l[valid]
  m <- m[valid]
  yr <- yr[valid]

  plot_data[[group]] <- data.frame(l = l, m = m, yr = yr, Group = group)
  label_data[[group]] <- plot_data[[group]] %>% filter(yr %% 3 == 0)

  fit <- NULL
  tryCatch({
    fit <- nls(m ~ (a * l) / (1 + b * l),
               start = list(a = max(m), b = 1 / max(l)),
               data = data.frame(l, m))

    l_pred <- seq(min(l), max(l), length.out = 100)
    m_pred <- predict(fit, newdata = data.frame(l = l_pred))
    curve_data[[group]] <- data.frame(l = l_pred, m = m_pred, Group = group)

    coefs <- coef(fit)
    params <- rbind(params, data.frame(Model = group, a = coefs["a"], b = coefs["b"]))

  }, error = function(e) {
    message(paste("Model", group, ": Beverton-Holt fit failed ->", e$message))
  })
}

# Unir data frames
plot_data_all <- bind_rows(plot_data)
curve_data_all <- bind_rows(curve_data)
label_data_all <- bind_rows(label_data)

# Graficar
srr_sce <- ggplot() +
  geom_point(data = plot_data_all, aes(x = l, y = m, shape = Group), size = 2) +
  geom_line(data = curve_data_all, aes(x = l, y = m, color = Group), linewidth = 1.2) +
  ggrepel::geom_text_repel(data = label_data_all, aes(x = l, y = m, label = yr, color = Group),
                           size = 4, max.overlaps = 50) +
  scale_shape_manual(values = shapes) +
  scale_color_manual(values = colors) +
  theme_bw() +
  ylim(0, max(plot_data_all$m)) +
  labs(x = "Spawning Stock Biomass (SSB)", y = "Predicted Recruitment", shape = "Model", color = "Model") +
  theme(legend.position = "bottom")
```

### Productivity and Interannual Variability by Scenario

Estimating the productivity of Antarctic krill is critical for understanding the species’ capacity to replenish its population in response to varying levels of spawning biomass. Productivity, defined as the ratio of recruitment to spawning stock biomass, provides a standardized measure of reproductive success and population resilience under different ecological and fishing pressures. Comparing productivity across scenarios—each representing different assumptions about environmental drivers, fishing mortality, or predator dynamics—enables a robust evaluation of how krill populations reflect this changes.

For each scenario \( i \) and year \( t \), we computed the productivity as the ratio between recruitment and spawning stock biomass (SSB):

$$
\text{Productivity}_{i,t} = \frac{\text{Recruitment}_{i,t}}{\text{SSB}_{i,t}}
$$

We also calculated the **interannual percentage change** in recruitment and SSB as:

$$
\text{Change in Recruitment}_{i,t} = \left( \frac{\text{Recruitment}_{i,t} - \text{Recruitment}_{i,t-1}}{\text{Recruitment}_{i,t-1}} \right) \times 100
$$

$$
\text{Change in SSB}_{i,t} = \left( \frac{\text{SSB}_{i,t} - \text{SSB}_{i,t-1}}{\text{SSB}_{i,t-1}} \right) \times 100
$$

These metrics allow us to analyze both the productivity and the temporal dynamics of the population under each scenario \( i \).

```{r}
# Simular 4 dataframes por escenario (usa tus datos reales)
escenarios <- list(
  s1.1 = base.model1.1$SPAWN_RECR_CURVE,
  s1.2 = base.model1.2$SPAWN_RECR_CURVE,
  s1.3 = base.model1.3$SPAWN_RECR_CURVE,
  s1.4 = base.model1.4$SPAWN_RECR_CURVE
)

# Crea una tabla con productividad y cambios para cada escenario
resultados <- purrr::imap_dfr(escenarios, function(df, nombre) {
  df %>%
    mutate(
      Scenario = nombre,
      Productividad = Recruitment / SSB,
      Cambio_Recruit = c(NA, diff(Recruitment)) / lag(Recruitment) * 100,
      Cambio_SSB = c(NA, diff(SSB)) / lag(SSB) * 100
    )
})

prod <- ggplot(resultados, aes(x = `SSB/SSB_virgin`, y = Productividad, 
                       color = Scenario)) +
  geom_point(size = 1.1) +
  labs(title = "",
       y = "Recruitment / SSB",
       x = "SSB / SSB_virgin") +
  scale_color_manual(name= "",
                    values = c("s1.1" = "#ca0020", 
                               "s1.2" = "#f4a582", 
                               "s1.3" = "#bababa", 
                               "s1.4" = "#404040")) +
  theme_bw() +
  theme(legend.position = "bottom")
```

These two panels provide insight into the stock-recruitment dynamics of Antarctic krill under four assessment scenarios, each incorporating different levels of ecosystem complexity.

**Left panel – Predicted Recruitment vs Spawning Stock Biomass (SSB):**  
This plot compares the predicted recruitment across years with corresponding spawning biomass values for the four model scenarios (s1.1 to s1.4). Scenario s1.1 (red) shows a relatively flat and optimistic Beverton-Holt relationship, with high recruitment predicted even at low levels of SSB. The other scenarios, particularly s1.3 (gray) and s1.4 (black), show much more constrained recruitment estimates and a more saturating relationship, where recruitment increases only slightly with increasing SSB. This suggests that incorporating environmental (s1.3) and predator (s1.4) effects leads to a more conservative and ecologically realistic recruitment dynamic. Year labels illustrate how certain years (e.g., 2004, 2016, 2022) deviate between scenarios, especially at low biomass levels.

**Right panel – Recruitment per SSB vs Relative SSB (SSB / SSBo):**  
This plot displays recruitment efficiency (Recruitment/SSB) as a function of relative spawning biomass (SSB / SSB₀). In all models, recruitment efficiency is highest at very low biomass levels and declines as SSB increases, consistent with compensatory dynamics. However, s1.3 and s1.4 (which include environmental effects) show significantly lower overall productivity across all SSB levels compared to s1.1 and s1.2. This indicates that the inclusion of ecosystem drivers reduces the per-capita productivity of the stock, especially under depleted conditions.

**Interpretation:**  
Together, these plots indicate that ecosystem-informed models (s1.3 and s1.4) produce more constrained and precautionary estimates of krill productivity. The more optimistic recruitment predicted by s1.1 (and to a lesser extent s1.2) may overestimate stock resilience, particularly in low SSB conditions. These findings emphasize the importance of accounting for environmental variability and predator dynamics in stock assessment models to avoid overoptimistic management advice and better reflect the biological limits of the krill population.

```{r, fig.height=4, fig.width=6, fig.cap="Left panel: Stock–recruitment relationships for Antarctic krill under four assessment scenarios. Predicted recruitment as a function of spawning stock biomass (SSB), with fitted Beverton-Holt curves and annual labels for selected years. Right panel: Recruitment per unit of SSB plotted against relative SSB (SSB/SSBo), illustrating differences in per-capita productivity across models ( s1.1 (no ecosystem variables), s1.2 (with predator effects), s1.3 (with environmental variables), and s1.4 (with both predator and environmental effects))"}
ggarrange(srr_sce,
          prod,
          ncol = 2)
```

### Compara Fishing Mortality.

```{r eval=FALSE}
escenarios <- list(
  s1.1 = base.model1.1$exploitation[1:5],
  s1.2 = base.model1.2$exploitation[1:5],
  s1.3 = base.model1.3$exploitation[1:5],
  s1.4 = base.model1.4$exploitation[1:5]
)

# Calcula el cambio porcentual entre años para F_std y annual_F
# Calcula el cambio relativo entre años para F_std y annual_F
# Calcula el cambio relativo entre años para F_std y annual_F
resultados <- purrr::imap_dfr(escenarios, function(df, nombre) {
  df %>%
    mutate(
      Scenario = nombre,
      Relativo_Fstd = c(NA, diff(F_std) / head(F_std, -1)),  # Cambio relativo de F_std
      Relativo_annual_F = c(NA, diff(annual_F) / head(annual_F, -1))  # Cambio relativo de annual_F
    )
})

# Visualiza los resultados en un gráfico
ggplot(resultados, aes(x = Yr, y = Relativo_Fstd, color = Scenario)) +
  geom_line(size = 1.1) +
  labs(title = "Cambio porcentual en F_std entre años por escenario",
       x = "Año",
       y = "Cambio porcentual en F_std") +
  scale_color_manual(name= "Scenario",
                    values = c("s1.1" = "#ca0020", 
                               "s1.2" = "#f4a582", 
                               "s1.3" = "#bababa", 
                               "s1.4" = "#404040")) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")

# Gráfico para annual_F
ggplot(resultados, aes(x = Yr, y = Relativo_annual_F, color = Scenario)) +
  geom_line(size = 1.1) +
  labs(title = "Cambio porcentual en annual_F entre años por escenario",
       x = "Año",
       y = "Cambio porcentual en annual_F") +
  scale_color_manual(name= "Scenario",
                    values = c("s1.1" = "#ca0020", 
                               "s1.2" = "#f4a582", 
                               "s1.3" = "#bababa", 
                               "s1.4" = "#404040")) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")

@@

### AKL
